\textsc{Atomic modeling of argon}\\ \footnotesize{\textsc{Project 5 --- FYS3150}}

### 2015-11-28 projects/project_5/molecular_dynamics/article/article.tex
working on energy fluctuations

    \begin{enumerate}[1)]
        \item Introduce periodic boundary conditions.
        \item Implement a function for setting the net-momentum for the system to zero.
        \item Make the atoms conform to a face-centered cubic lattice configuration.
        \item Introduce the force model to the system, given by the Lennard-Jones potential.
        \item Implement the Velocity-Verlet algorithm in order to get a better
            conservation of energy.
        \item Write functions for sampling and storing physical quantities
    \end{enumerate}

    Running the program in its initial state yields a random configuration of
    the atoms. Since there is no force acting on the molecules they move in
    straight lines determined by their initial velocity vector. Their motion is
    unbounded, due to the periodic boundary condition not being implemented
    yet. This can be seen in figure \ref{fig:first_iteration}.
    \begin{figure}
        \centering \includegraphics[width=0.4\linewidth]{first_iteration.png}
        \caption[Initial state of the program]{First iteration of the program --- no boundary conditions and
            no force acting on the atoms. The atoms move with a constant speed
            in the direction of their initial velocity and the initial
        positions are uniformly distributed.}
        \label{fig:first_iteration}
    \end{figure}

@@

@@

We initially decided to use the Velocity-Verlet integration method instead of
Euler-Cromer because of its good energy conservation. It is therefore of
interest to examine exactly what kind of impact this change of integration
method has for our simulations.  We now simulate a set of similar systems with
varying temperature and look at how the fluctuations in total energy scales
with the timestep $\Delta T$. In order to do this we can consider the
standard-devation of the total energy.

\subsection{Melting temperature of system}
\label{sub:melting_temperature_of_system}

>>>>

    \begin{enumerate}[1)]
        \item Introduce periodic boundary conditions.
        \item Implement a function for setting the net-momentum for the system to zero.
        \item Make the atoms conform to a face-centered cubic lattice configuration.
        \item Introduce the force model to the system, given by the Lennard-Jones potential.
        \item Implement the Velocity-Verlet algorithm in order to get a better
            conservation of energy.
        \item Write functions for sampling and storing physical quantities.
        \item Miscellaneous changes.
    \end{enumerate}

    Running the program in its initial state yields a random configuration of
    the atoms. Since there is no force acting on the molecules they move in
    straight lines determined by their initial velocity vector. Their motion is
    unbounded, due to the periodic boundary condition not being implemented
    yet. This can be seen in figure \ref{fig:first_iteration}.

    \begin{figure}
        \centering \includegraphics[width=0.4\linewidth]{first_iteration.png}
        \caption[Initial state of the program]{First iteration of the program --- no boundary conditions and
            no force acting on the atoms. The atoms move with a constant speed
            in the direction of their initial velocity and the initial
        positions are uniformly distributed.}
        \label{fig:first_iteration}
    \end{figure}

@@

    \subsection{Miscellaneous changes}
    \label{sub:miscellaneous_changes}

    In order to ease the implementation of the python framework used to run the
    simulations, we need to make some auxiliary changes to the
    \texttt{main}-method of the program. First of all, we are interested in
    comparing the two integration methods the program currently supports, so we
    add a fourth possible command line argument for setting the integration
    method. The program defaults to Velocity-Verlet.

    We also need a way of changing the time step $\Delta t$ used in the
    simulations. This is also implemented as a command line argument, and we
    use a default time step of $\Delta t = 1.0 \times 10^{-15}$ seconds.

@@

We initially decided to use the Velocity-Verlet integration method instead of
Euler-Cromer because of its good energy conservation. It is therefore of
interest to examine exactly what kind of impact this change of integration
method has for our simulations.  We now simulate a set of similar systems with
varying temperature and look at how the fluctuations in total energy scales
with the time step $\Delta T$. In order to do this we can consider the
standard-deviation of the total energy.

We are going to run the simulations for a starting temperature of 300 kelvin,
and vary the time step from $1.0 \times 10^{-16}$ to $1.0 \times 10^{-13}$
seconds.
\subsection{Melting temperature of system}
\label{sub:melting_temperature_of_system}
